---
  title: "BIOS 611 HW2 Data transformation (Chapter 5)"
  author: "Enying Gao, 720464447"
  date: "`r format(Sys.time(), '%m/%d/%Y')`"
  output: html_document
---
##### Due Date: 09/10/18 (6pm)

  This set of exercise is taken from R for Data Science by Garrett Grolemund and Hadley Wickham but modified slightly for the purpose of this class.

# Exercise 1

Note that the dataset `flights` is from the package `nycflights13`

1.  Find all flights that (Shows all the rows with the following characteristics. Then for each problem, make a variable called 'ans.flight.i' where 'i' is replaced by the subproblem number(1,2,3,...) containing all these rows.)

```{r, include = FALSE}
library(nycflights13)
library(tidyverse)
```

    1. Had an arrival delay of two or more hours
    
```{r}
ans.flight.1 <- filter(flights, arr_delay >= 120)
```

    2. Flew to Houston (`IAH` or `HOU`)
    
```{r}
ans.flight.2 <- filter(flights, dest == 'IAH' | dest == 'HOU')
```

    3. Were operated by United, American, or Delta
    
```{r}
ans.flight.3 <- filter(flights, carrier %in% c('UA', 'AA', 'DL'))
```

    4. Departed in summer (July, August, and September)
    
```{r}
ans.flight.4 <- filter(flights, month %in% c(7, 8, 9))
```

    5. Arrived more than two hours late, but didn't leave late
    
```{r}
ans.flight.5 <- filter(flights, arr_delay > 120, dep_delay <= 0)
```

    6. Were delayed by at least an hour, but made up over 30 minutes in flight
    
```{r}
ans.flight.6 <- filter(flights, dep_delay >= 60, arr_delay-dep_delay < 30)
```

    7. Departed between midnight and 6am (inclusive)

```{r}
ans.flight.7 <- filter(flights, dep_time <= 600 | dep_time == 2400)
```

2.  Another useful dplyr filtering helper is `between()`. What does it do?
    Can you use it to simplify the code needed to answer the previous 
    challenges? (Write the code using between() for subproblem 7 in the previous problem)
    
    It simplifies the code for selecting observations within a range of values.

```{r}
ans.flight.7.1 <- filter(flights, !between(dep_time, 601, 2359))
```

3.  How many flights have a missing `dep_time`? (Make a varible called "num.flights.missing.dep" for this number) What other variables are missing? (Make a variable called "missing.var" containing the name of these missing varaibles) What might these rows represent?  
  
```{r}
num.flights.missing.dep <- nrow(filter(flights, is.na(dep_time)))
test <- colSums(is.na(flights)) > 0
y <- TRUE
missing.var <- names(test[test %in% y])
```
    
    8255 flights have a missing`dep_time`.
    
    These rows might represent flights that got cancelled.
    
4.  Why is `NA ^ 0` not missing? Why is `NA | TRUE` not missing?
    Why is `FALSE & NA` not missing? Can you figure out the general
    rule?  (`NA * 0` is a tricky counterexample!)   
      
    (Hint: Think about what NA means and think logically why the expressions above can be evaluated.)
    
    `NA ^ 0` is not missing because `NA` is treated as a number here. `NA^0` = 1.

    `NA | TRUE` is not missing because condition `or TRUE` is in the statement, which makes it always a `TRUE` statement.

    `FALSE & NA` is not missing because condition `and FALSE` is in the statement, which makes it always a `FALSE` statement.

    `NA * 0` is missing because `NA` can be any number. It can be a  infinite too, which would yield NA when doing multiplication.

```{r}
NA ^ 0
NA | TRUE
FALSE & NA
NA * 0
```

# Exercise 2

1.  How could you use `arrange()` to sort all missing values to the start?
    (Hint: use `is.na()`).  Make a variable called "missing.first" for this sorted tibble.
    
```{r}
missing.first <- arrange(flights, desc(is.na(dep_delay)))
```

2.  Sort `flights` to find the most delayed flights.(Make a variable called "delayed.sorted" for this sorted tibble).

```{r}
delayed.sorted <- arrange(flights, desc(dep_delay))
```

3. Sort `flights` to find the fastest flights (Make a variable called "fastest.flights" containing the rows with the fastest flights in terms of average speed).

```{r}
fastest.flights <- arrange(flights, desc(distance/air_time))[1,]
```

4.  Which flights travelled the longest? Which travelled the shortest? (Make a variable called "longest.flight" and "shortest.flight" for the rows with the longest and shortest flight)

```{r}
longest.flight <- arrange(flights, desc(distance))[1,]
shortest.flight <- arrange(flights, distance)[1,]
```

# Exercise 3

1.Brainstorm as many ways as possible to select `dep_time`, `dep_delay`,
    `arr_time`, and `arr_delay` from `flights`.
    
```{r}
select(flights, dep_time, dep_delay, arr_time, arr_delay)
select(flights, dep_time:arr_delay)
select(flights, starts_with('dep'), starts_with('arr'))
select(flights, dep_time, arr_time, ends_with('delay'))
```

2.  What happens if you include the name of a variable multiple times in
    a `select()` call?  
    
```{r}
select(flights, year, year)
```

    The `select` statement will ignore the times the name of a varibale appears in the code.

3.  What does the `one_of()` function do? Why might it be helpful in conjunction with this vector?
    
    ```{r}
    vars <- c("year", "month", "day", "dep_delay", "arr_delay")
    select(flights, one_of(vars))
    ```
    
    `one_of()` selects any variables within the arguments. In conjunction with this vector, we can extract these specific columns in flights.

# Exercise 4

1.  Currently `dep_time` and `sched_dep_time` are convenient to look at, but
    hard to compute with because they're not really continuous numbers. 
    Convert them to a more convenient representation of number of minutes
    since midnight. (Replace the variable "flights" with this tibble with converted dep_time and sched_dep_time)
    
```{r}
transmute(flights,
          dep_time_converted = dep_time %/% 100 * 60 + dep_time %% 100,
          sched_dep_time_converted = sched_dep_time %/% 100 * 60 + sched_dep_time %% 100)
```

2.  Compare `air_time` with `arr_time - dep_time`. What do you expect to see?
    What do you see? What do you need to do to fix it?
      
```{r}
transmute(flights, air_time, arr_time - dep_time)
transmute(flights, air_time,
          difference_in_arr_time_dep_time = arr_time %/% 100 * 60 + arr_time %% 100 -
            (dep_time %/% 100 * 60 + dep_time %% 100))

```

    I expect them to be the same.
    
    However, they're different when I run the codes, because arr_time and dep_time are not conatinuous numbers.
    
    I will change arr_time and dep_time to minutes first.
    
    The resulting values are still not the same.

3.  Compare `dep_time`, `sched_dep_time`, and `dep_delay`. How would you
    expect those three numbers to be related?

```{r}
select(flights, dep_time, sched_dep_time, dep_delay)
```

    I expect `dep_delay` to be the difference between `dep_time` and `sched_dep_time`. `dep_delay` is positive if flights were delayed. `dep_delay` is negative if flights left early than scheduled.

4.  Find the 10 most delayed flights using a ranking function.(Name this variable "top.delayed" containing the rows) How do you want 
    to handle ties? Carefully read the documentation for `min_rank()`.

```{r}
top.delayed <- arrange(flights, min_rank(desc(dep_delay)))[1:10,]
```

    `min_rank()` uses `ties.method = 'min'` to handle ties, meaning ties will receive the same rank.

5.  What does `1:3 + 1:10` return? Why?  
    
```{r}
1:3 + 1:10
```

    It returns a warning message saying that `longer object length is not a multiple of shorter object length`. Since `1:3` has a shorter length than `1:10`, `1:3` will be automatically extended so that the argument can be calculated. Meaning that the 4th element in the reuslt is calculated as 1 + 4 = 5.

# Exercise 5

1.  Brainstorm at least 3 different ways to assess the typical delay 
    characteristics of a group of flights. Consider the following scenarios:
    
A flight is 15 minutes early 50% of the time, and 15 minutes late 50% of the time.

```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))

# When referring to late, I assume the question is talking about arrival delay.

# A flight is at least 15 minutes early 50% of the time, and at least 15 minutes late 50% of the time.
not_cancelled %>% 
  group_by(flight) %>% 
  summarise(minutes_early = sum(arr_delay <= -15)/n(),
            minutes_late = sum(arr_delay >= 15)/n()) %>% 
  filter(minutes_early == 0.5, minutes_late == 0.5)
```
      
A flight is always 10 minutes late.

```{r}
# Always means greater than 90% of the time.
not_cancelled %>% 
  group_by(flight) %>% 
  summarise(percentage = sum(arr_delay >= 10)/n()) %>% 
  filter(percentage > 0.9)
```
    
A flight is 30 minutes early 50% of the time, and 30 minutes late 50% of the time.

```{r}
# A flight is at least 30 minutes early 50% of the time, and at least 30 minutes late 50% of the time.
not_cancelled %>% 
  group_by(flight) %>% 
  summarise(minutes_early = sum(arr_delay <= -30)/n(),
            minutes_late = sum(arr_delay >= 30)/n()) %>% 
  filter(minutes_early == 0.5, minutes_late == 0.5)
```
      
99% of the time a flight is on time. 1% of the time it's 2 hours late.

```{r}
not_cancelled %>% 
  group_by(flight) %>% 
  summarise(on_time = sum(arr_delay <= 0)/n(),
            late = sum(arr_delay >= 120)/n()) %>% 
  filter(on_time >= 0.99, late <= 0.01)
```
    
Which is more important: arrival delay or departure delay?
    
    Arrival delay is more important because even though departure delay can be greater than 0, a flight can catch up the time in the air, causing no influence to arrival time.
    
(Make 3 similar statements as the ones above using the data we have)
  
    Over 80% of the time a flight travles over 2500 miles.
    
```{r}
not_cancelled %>% 
  group_by(flight) %>% 
  summarise(percentage = sum(distance > 2500)/n()) %>% 
  filter(percentage > 0.8)
```
    
    95% of the time a flight has no departure delays.
    
```{r}
not_cancelled %>% 
  group_by(flight) %>% 
  summarise(percentage = sum(dep_delay <= 0)/n()) %>% 
  filter(percentage > 0.95)
```
    
    93% of the time a flight arrives less than 15 minutes late.
    
```{r}
not_cancelled %>% 
  group_by(flight) %>% 
  summarise(percentage = sum(arr_delay <15)/n()) %>% 
  filter(percentage > 0.93)
```
    
    
```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))

not_cancelled %>% 
  group_by(flight) %>% 
  summarise(minutes_early = sum(arr_delay == -15)/n(),
            minutes_late = sum(arr_delay == 15)/n()) %>% 
  filter(minutes_early == 0.5, minutes_late == 0.5)

not_cancelled %>% 
  group_by(flight) %>% 
  summarise(minutes_early = sum(arr_delay == -30)/n(),
            minutes_late = sum(arr_delay == 30)/n()) %>% 
  filter(minutes_early == 0.5, minutes_late == 0.5)


```

2.  Look at the number of cancelled flights per day. Is there a pattern?
    Is the proportion of cancelled flights related to the average delay?


```{r}
# Removing all cancelled flights.
flights %>% 
  filter(is.na(dep_delay), is.na(arr_delay)) %>% 
  group_by(year, month, day) %>% 
  summarise(total = n())

flights %>% 
  group_by(year, month, day) %>% 
  summarise(prop = sum(is.na(dep_delay))/n(),
            average = mean(arr_delay, na.rm = TRUE)) %>% 
  filter(prop < 0.1) %>% 
  ggplot(mapping = aes(x = average, y = prop)) + 
    geom_point(alpha = 1/5) +
    geom_smooth(se=FALSE)
```

    There seems to be a positive relationship between proportion of cancelled flights and the average delay. As proportion of cancelled flights increases, average delay increases.

3. Which carrier has the worst delays? 
  
```{r}
flights %>% 
  group_by(carrier) %>% 
  summarise(avg = mean(arr_delay, na.rm = TRUE)) %>% 
  arrange(desc(avg))
```

    In terms of percentage of delayed arrivals, F9(Frontier Airlines) is the worst.

4.  What does the `sort` argument to `count()` do. When might you use it?  
  
    It sorts `count()` in descending order. It helps to arrange the number of occurences from biggest to smallest.

# Exercise 6

1.  Which plane (`tailnum`) has the worst on-time record? (Make a variable called "worst.plane" for this plane tail number)

```{r}
worst.plane <-
  flights %>% 
  filter(arr_delay > 0) %>% 
  group_by(tailnum) %>%
  summarise(avg = mean(arr_delay)) %>% 
  arrange(desc(avg))
```


2. For each destination, compute the total minutes of delay. (Make a tibble called "dest.delayedmin" with these two columns) 
    
```{r}
dest.delayedmin <- 
  flights %>% 
  group_by(dest) %>% 
  filter(arr_delay > 0) %>% 
  summarise(total = sum(arr_delay))
```

3.  Delays are typically temporally correlated: even once the problem that
    caused the initial delay has been resolved, later flights are delayed 
    to allow earlier flights to leave. Using `lag()`, explore how the delay
    of a flight is related to the delay of the immediately preceding flight.
(What proportion of delayed flights is caused by the delay of the immediately preceding flights)  
  
```{r}
# Removing all cancelled flights.
flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))  %>%
  mutate(pre_delay = lag(dep_delay)) %>% 
  ggplot(mapping = aes(x = dep_delay, y = pre_delay)) +
  geom_point(alpha = 1/10) +
  geom_smooth(se = FALSE)

flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))  %>%
  mutate(pre_delay = lag(dep_delay)) %>% 
  summarise(percentage = sum(dep_delay > 0 & pre_delay > 0, na.rm = TRUE) /
            sum(dep_delay > 0, na.rm = TRUE))
```
    
    The plot suggests that the delay of a flight is positively related to the delay of the immediately preceding flight, meaning. In other words, if a flight is delayed, then it's very likely that the immediately preceding flight is delayed too.
    
    In terms of proportion, about 52% of delayed flights is caused by the delay of teh immediately preceding flights.

4. Find all destinations that are flown by at least two carriers. (Put these destinations into a variable called "popular.destinations")

```{r}
# Removing all cancelled flights.
popular.destinations <- 
  flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))  %>% 
  group_by(dest) %>% 
  summarise(carriers = n_distinct(carrier)) %>% 
  filter(carriers >= 2)
```

5.  For each plane, count the number of flights before the first delay of greater than 1 hour. (Make a tibble called "plane.num.before.first.delay" with these two columns)
    
```{r}
# Removing all cancelled flights.
plane.num.before.first.delay <-
  flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))  %>%
  group_by(tailnum) %>% 
  summarise(first = sum(cummax(dep_delay) < 60)) %>% 
  arrange(desc(first))
```
