---
  title: "BIOS 611 HW6 Functions, vectors, and iteration (Chapter 19 - 21)"
  author: "Enying Gao"
  date: "`r format(Sys.time(), '%m/%d/%Y')`"
  output: html_document
---
  (This homework is due Mon. Oct. 15th at 6 PM.)

  This set of exercise is largely taken from R for Data Science by Garrett Grolemund and Hadley Wickham.

```{r, include=FALSE}
library(tidyverse)
library(stringr)
```


# Exercise 1

1.  `dim()` function only works for objects such as matrix, array, or data frame. It returns `NULL` when you apply the function to a vector or a list.
    Write a function called `dim2()` that returns dimension of an object, where the dimension is defined by `dim()` for matrices, arrays, and data frames, by `length()` for vectors and lists, and by `NULL` for all other classes.  
    Hint: use `if () {} else {}` statement.
    
    Answer: 

    ```{r}
    dim2 <- function(x) {
      if (is.null(dim(x))) {
        length(x)
      } else {
        dim(x)
      }
    }
    ```

2.  Write a function that returns the number of `.R` files of a specific folder (input: `path`). 
    Hint: use `list.files()` and `grep()` (or `str_detect()`).  
    (For those code chunks with `eval = FALSE` option, please leave the eval option as it is, 
     because the output will not be reproducible.)
    
    Answer: 

    ```{r, eval = FALSE}
    number_of_R_files <- function(path) {
      sum(str_detect(list.files(path), ".R$"))
    }
    ```


3.  The following function was designed to convert the existing count data to a compositional data.  
    (i.e. each column is divided by its column sum.)
    The code has several problems. Criticize the code and provide a solution.

    ```{r}
    notGood <- function(mat, margin) {
      tmp <- apply(mat, 2, sum)
      mat <- mat / tmp
    }
    
    # example data
    set.seed(1)
    dat <- matrix(rpois(100, rep(1:10, each = 10)), 10)
    ```

    Answer: 

    The margin is not specified inside the function, tmp is supposed to be margin.

    ```{r}
    Good <- function(dat) {
      margin <- apply(dat, 2, sum)
      dat <- dat / margin
      return(dat)
    }
    ```


# Exercise 2

1.  What's the difference between `if` and `ifelse()`? Carefully read the help
    and construct three examples that illustrate the key differences.
    Keywords: type of outputs, `NA`, output for `FALSE`
    
    Answer: 

    ```{r, eval = FALSE}
    # Example 1: Type of outputs
    x = 3
    if(x > 2) {y = "Yes"}
    ifelse(x > 2, y = "Yes", y = "No")
    
    # Example 2: `NA`
    x <- c("NA")
    if(x > 2) {y = "Yes"}
    ifelse(x > 2, y = "Yes", y = "No")
    
    # Example 3: Output for `FALSE`
    x = 1
    if(x > 2) {y = "Yes"}
    ifelse(x > 2, y = "Yes", y = "No")
    ```

  `if` evaluates condition that has length 1, while `ifelse` evalutates every elements in the condition. When encountering `NA`, `if` statement doesn't respond anything, while `ifelse` statement returns an error. With respect to output for `FALSE`, `if` returns no value for false condition while `ifelse` returns value we specify for the `False` condition.

2.  How could you use `cut()` to simplify this set of nested if-else statements?

    ```{r, eval = FALSE}
    if (temp <= 0) {
      "freezing"
    } else if (temp <= 10) {
      "cold"
    } else if (temp <= 20) {
      "cool"
    } else if (temp <= 30) {
      "warm"
    } else {
      "hot"
    }
    ```
    
    How would you change the call to `cut()` if I'd used `<` instead of `<=`?
    What is the other chief advantage of `cut()` for this problem? (Hint:
    what happens if you have many values in `temp`?)

    Answer: 

    ```{r, eval = FALSE}
    cut(temp, c(-Inf, 0, 10, 20, 30, Inf), right = FALSE,
    labels = c("freezing", "cold", "cool", "warm", "hot"))
    ```


3.  What does this `switch()` call do? What happens if `x` is "e"?

    ```{r, eval = FALSE}
    switch(x, 
      a = ,
      b = "ab",
      c = ,
      d = "cd"
    )
    ```
    
    Experiment, then carefully read the documentation. 

    Answer: 

    ```{r}
    switch("a", 
      a = ,
      b = "ab",
      c = ,
      d = "cd"
    )
    switch("b", 
      a = ,
      b = "ab",
      c = ,
      d = "cd"
    )
    switch("e", 
      a = ,
      b = "ab",
      c = ,
      d = "cd"
    )
    ```

    `switch` evaluates expression and return values based on the arguments. If `switch` encounters an arguemnt with missing value, it will take the next value that is not missing. If the expression in `switch` (i.e "e") doesn't match any arguemnts in the statemnet, it will return nothing as there's no specification on what to return for null value.

# Exercise 3

1.  What does `commas(letters, collapse = "-")` do? Why?

    Answer: 

    ```{r}

    ```

    It combines letters together with "-".

2.  The default value for the `method` argument to `cor()` is 
    `c("pearson", "kendall", "spearman")`. What does that mean? What 
    value is used by default?

    Answer: 
    
    Depends on the method we choose, we can compute correlation coefficient differently. The default method is pearson.


# Exercise 4

1.  Describe the difference between `is.finite(x)` and  `!is.infinite(x)`.

    Answer: 

    ```{r}
    is.finite(1)
    !is.infinite(1)
    is.finite(Inf)
    !is.infinite(Inf)
    is.finite(NA)
    !is.infinite(NA)
    is.finite(NaN)
    !is.infinite(NaN)
    ```

    The only difference is how `is.finite(x)` and `!is.infinite(x)` treat NA and NaN. The results are opposite for these two functions.

2.  Create functions that take a vector as input and returns:
    
   1. The last value.  Should you use `[` or `[[`?

    Answer: 

    ```{r}
    x <- c(a = 10, b = 20)
    function(x) {
      x[[length(x)]]
    }
    ```

    We should use `[[` because we want to extract a single element

    1. The elements at even numbered positions.
    
    Answer: 

    ```{r}
    function(x) {
      x[seq_along(x) %% 2 == 0]
      }
    ```

    1. Every element except the last value.
    
    Answer: 

    ```{r}
   function(x) {
      x[-length(x)]
      }
    ```

3.  What happens when you subset with a positive integer that's bigger
    than the length of the vector? What happens when you subset with a 
    name that doesn't exist?

    Answer: 

    ```{r}
    x <- c(y = 2, z = 4)
    x[3]
    x["w"]
    ```

    The results will both be `NA`.

# Exercise 5

1.  Describe the difference between `abc[[1]]` and `abc[["1"]]`.  
    Hint: You can use `abc <- list("2"=1:3, "1"=4:6)`

    Answer: 

    ```{r}
    abc <- list("2"=1:3, "1"=4:6)
    abc[[1]]
    abc[["1"]]
    ```

    `abc[[1]]` extracts the first component from list abc, while `abc[["1"]]` extracts the component that's named "1" from list abc.
    

2.  What does `hms::hms(3600)` return? How does it print? What primitive
    type is the augmented vector built on top of? What attributes does it 
    use?
    
    Answer: 

    ```{r}
    hms::hms(3600)
    typeof(hms::hms(3600))
    attributes(hms::hms(3600))
    ```

    It returns a time and prints in terms of hour, minutes and seconds. The augmented vector is built on top of double and it uses attributes class and units.

3.  Try and make a tibble that has columns with different lengths. What
    happens?

    Answer: 

    ```{r, eval = FALSE}
    tibble(x = 1:5, y = 5:3)
    ```

    It gives me an error, telling me that y must be length 1 or 5.

4.  Based on the definition above, is it ok to have a list as a
    column of a tibble? Provide an example, if yes.
    
    Answer: 

    ```{r}
    tibble(x = list(1:5), y = 5:3)
    ```

    Yes, it is ok to have a list as a column of a tibble.


# Exercise 6

1.  Write for loops to:


    1. Compute the mean of every column in `mtcars`.
    
    Answer: 

    ```{r}
    output <- vector("double", ncol(mtcars))
    for (i in seq_along(mtcars)) {           
      output[[i]] <- mean(mtcars[[i]])  
      }
    output
    ```

    1. Determine the type of each column in `nycflights13::flights`.
    
    Answer: 

    ```{r}
    output <- vector("list", ncol(nycflights13::flights))
    for (i in seq_along(nycflights13::flights)) {           
      output[[i]] <- class(nycflights13::flights[[i]])  
      }
    output
    ```

    1. Compute the number of unique values in each column of `iris`.
    
    Answer: 

    ```{r}
    output <- vector("double", ncol(iris))
    for (i in seq_along(iris)) {           
      output[[i]] <- length(unique(iris[[i]]))
      }
    output
    ```

    1. Generate 10 random normals for each of $\mu = -10$, $0$, $10$, and $100$.
    
    Think about the output, sequence, and body __before__ you start writing
    the loop.

    Answer: 

    ```{r}
    mu <- c(-10, 0, 10, 100)
    output <- vector("list", length(mu))
    for (i in seq_along(mu)) {           
      output[[i]] <- rnorm(10, mu[i])
      }
    output
    ```

2.  Eliminate the for loop in each of the following examples by taking advantage of an existing function that works with vectors:
    (Do not use `apply`, `lapply`, `do.call` kind of functions, but try to use very basic functions)
    
    ```{r, eval = FALSE}
    # a.
    out <- ""
    for (x in letters) {
      out <- stringr::str_c(out, x)
    }
    
    # b.
    x <- sample(100)
    sd <- 0
    for (i in seq_along(x)) {
      sd <- sd + (x[i] - mean(x)) ^ 2
    }
    sd <- sqrt(sd / (length(x) - 1))
    
    # c.
    x <- runif(100)
    out <- vector("numeric", length(x))
    out[1] <- x[1]
    for (i in 2:length(x)) {
      out[i] <- out[i - 1] + x[i]
    }
    ```

    Answer: 

    ```{r}
    # a.
    stringr::str_c(letters, collapse = "")
    
    # b.
    sd(x)
    
    # c.
    cumsum(x)
    ```


3.  It's common to see for loops that don't preallocate the output and instead
    increase the length of a vector at each step:
    
    ```{r, eval = FALSE}
    output <- vector("integer", 0)
    for (i in seq_along(x)) {
      output <- c(output, lengths(x[[i]]))
    }
    output
    ```
    
    How does this affect performance? Design and execute an experiment.

    Answer: 

    ```{r}
    library(microbenchmark)
    function1 <- function(x) {
      out <- vector("integer", 0)
      out[1] <- x[1]
       for (i in 2:length(x)) {
        out <- c(out, out[i - 1] + x[i])
       }
    }
    
     function2 <- function(x) {
       out <- vector("numeric", length(x))
      out[1] <- x[1]
       for (i in 2:length(x)) {
         out[i] <- out[i - 1] + x[i]
       }
     }
    
    microbenchmark(function1(runif(100)), times = 3)
    microbenchmark(function2(runif(100)), times = 3)
    ```

    The performance is slower if loops don't preallocate the output and instead increase the length of a vector at each step.

# Exercise 7

1.  Imagine you have a directory full of CSV files that you want to read in.
    You have their paths in a vector, 
    `files <- dir("data/", pattern = "\\.csv$", full.names = TRUE)`, and now
    want to read each one with `read_csv()`. Write the for loop that will 
    load them into a single data frame. 
    (For those code chunks with `eval = FALSE` option, please leave the eval option as it is, 
     because the output will not be reproducible.)
    
    Answer: 

    ```{r, eval = FALSE}
    output <- vector("list", length(files))
    for (i in seq_along(files)) {           
      output[[i]] <- read_csv(files[[i]])  
      }
    dplyr::bind_rows(output)
    ```


2.  Write a function that prints the mean of each numeric column in a data 
    frame, along with its name. For example, `show_mean(iris)` would print:
    
    ```{r, eval = FALSE}
    show_mean(iris)
    #> Sepal.Length: 5.84
    #> Sepal.Width:  3.06
    #> Petal.Length: 3.76
    #> Petal.Width:  1.20
    ```
    
    (Extra challenge: what function did I use to make sure that the numbers
    lined up nicely, even though the variable names had different lengths?)

    Answer: 

    ```{r}
    show_means <- function(x) {
      column <- vector("logical", length(x))
      for (i in seq_along(x)) {
        column[[i]] <- is.numeric(x[[i]])
        }
      x <- x[column]
      for (i in seq_along(x)) {
        cat(str_c(str_pad(
          str_c(colnames(x)[i], ":"), max(str_length(colnames(x))) + 2, side = "right"),
          round(mean(x[[i]]), 2)), fill = TRUE)
      }
    }
    show_means(iris)
    ```

3.  What does this code do? How does it work?

    ```{r, eval = FALSE}
    trans <- list( 
      disp = function(x) x * 0.0163871,
      am = function(x) {
        factor(x, labels = c("auto", "manual"))
      }
    )
    for (var in names(trans)) {
      mtcars[[var]] <- trans[[var]](mtcars[[var]])
    }
    ```

    Answer: 

    The code changes disp by multiplying it by 0.0163871. It changes am by converting it to a factor with lables "auto" and "manual". It then replaces the original values in mtcars data with the new values for column am and disp.
    
# Exercise 8

1.  Write code that uses one of the `map` functions to:

    1. Compute the mean of every column in `mtcars`.
    Answer: 

    ```{r,}
    map_dbl(mtcars, mean)
    ```

    1. Determine the type of each column in `nycflights13::flights`.
    Answer: 

    ```{r}
    map(nycflights13::flights, class)
    ```

    1. Compute the number of unique values in each column of `iris`.
    Answer: 

    ```{r}
    map_dbl(iris, ~length(unique(.)))
    ```

    1. Generate 10 random normals for each of $\mu = -10$, $0$, $10$, and $100$.

    Answer: 

    ```{r}
    map(c(-10, 0, 10, 100), rnorm, n = 10)
    ```

2.  What happens when you use the map functions on vectors that aren't lists?
    What does `map(1:5, runif)` do? Why?
    
    Answer: 

    ```{r} 
    map(1:5, runif)
    ```

    They should work properly on vectors that aren't lists too. `map(1:5, runif)` creates 5 lists that have vector length of 1 to 5 respectively.

3.  Rewrite `map(x, function(df) lm(mpg ~ wt, data = df))` to eliminate the 
    anonymous function. 

    Answer: 

    ```{r, eval = FALSE}
    map(x, ~ lm(mpg ~ wt, data = .))
    ```