---
title: "Reproduce me."
author: "Jimmy Kimmel"
date: "8/24/2018"
output: html_document
---

## Linear regression

### Assumptions

Standard analysis of linear regression is based on the following *four assumptions*:

1. Linear relation with fixed regressors

1. Independence

1. Constant variance

1. Mean zero error

These assumptions can be succinctly written as: $Y_i = X_i^T \beta + \epsilon_i$, $\epsilon_i \sim^{iid} D(0, \sigma^2)$.


(Hint - LaTex scripts: \$Y_i = X_i^T \\beta + \\epsilon_i\$, \$\\epsilon_i \\sim^{iid} D(0, \\sigma^2)\$)

### Example

Using `mtcars` dataset, we fit a linear regression in `R`.


```{r}
out = lm(mpg ~ disp + wt, data = mtcars)

```
```{r, echo=FALSE}
library(knitr)
kable(summary(out)$coef, digits=2, caption = "Regression table")
```

The variable `wt` has statistically significant association with `mpg`.

Next, we present the residual plot.

(Hint: plot(lm.object, which = 1))

```{r, echo = FALSE, fig.height = 3, fig.width = 3, fig.cap = "residual plot"}
plot(out, 1)
```

